### 2016-09-18 Rmd/analyze.Rmd

---
title: "Untitled"
author: "Ross Cunning, Erik B. Muller, Ruth D. Gates, Roger R. Nisbet"
date: "September 15, 2016"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r initialize, include=FALSE}
# Initialize SU function
synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
# Source default parameters and functions
sapply(list.files(path="R", pattern="*.R", full.names=T), source, .GlobalEnv)
```

```{r lightamp, fig.width=5, fig.height=5, eval=F}
# Import data from Marcelino et al. 2013
md <- read.csv("data/marcelino2013_fig2a.csv")
md$S.R <- md$density * cm2.m2 / CmolR.m2 * CmolS.cellS  # Convert units of symbiont density to CmolS/CmolR
md$A <- md$E + 1  # Convert "E"=excess light metric of Marcelino to "A"=amplification factor
# Plot data
plot(md$A ~ md$S.R, pch=c(21,23,22)[factor(md$spp)], bg=c("white", "black", "gray")[factor(md$spp)], 
     ylab="internal/external light", ylim=c(1,3.2),
     xlab="Symbiont density (CmolS/CmolR)",
     main="Light Amplification")
legend("topright", legend=levels(md$spp), pch=c(21,23,22), pt.bg=c("white", "black", "gray"), inset=0.02)
# Fit exponential decay model and plot fitted values
mod <- nls(A ~ x + y * exp(z * S.R), data=md, start=list(x=1, y=2, z=-3))
fit <- predict(mod, list(S.R=seq(0, 0.5, 0.01)))
lines(fit ~ seq(0, 0.5, 0.01))
# Get parameters of fit
formula(mod); coef(mod)
```

### Run 1
```{r run1}
# Set run time vector
time <- seq(1, 365, 0.1)  # if single model run, use time input

# Initialize environment
env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,0), X=c(0,0,0))

# Set parameters
pars <- def_pars()  # Get default parameters
#pars$rpars$gammaR <- 0.03  # Change any default parameters

# Run simulation
run <- run_coral(time=time, env=env, pars=pars)
```

##### Plot run 1 symbiont PI curve
```{r run1_PI, fig.height=1}
sym_PI(pars)
```

##### Plot run 1 output
```{r run1_plots}
plot_run(run)

```

#### Run 1 with no symbiont recycling
```{r runNR_plots}
parsNR <- pars
parsNR$spars$sigmaNS <- 0
parsNR$spars$sigmaCS <- 0
runNR <- run_coral(time=time, env=env, pars=parsNR)
plot_run(runNR)
with(runNR, {

})
```

#### Compare amount of recycled sym nitrogen between runs
```{r compareN}
proprNS <- run$S$rNS / (run$S$rNS + run$H$rhoN*run$H$H/run$S$S)
par(mfrow=c(2,1))
plot(time, proprNS)
proprNS <- runNR$S$rNS / (runNR$S$rNS + runNR$H$rhoN*runNR$H$H/runNR$S$S)
plot(time, proprNS)

proprCS <- run$S$rCS / (run$S$rCS + run$H$jCO2*run$H$H/run$S$S)
par(mfrow=c(2,1))
plot(time, proprCS)
proprCS <- runNR$S$rCS / (runNR$S$rCS + (runNR$H$rCH + runNR$H$jCO2)*runNR$H$H/runNR$S$S)
plot(time, proprCS)
```

### Run 2 - tolerant symbiont
```{r run2, echo=F}
pars2 <- pars
pars2$spars$kROS <- 80
pars2$spars$jCPm <- 1.4

run2 <- run_coral(time=time, env=env, pars=pars2)
```

##### Plot symbiont PI curve
```{r run2_PI, fig.height=2}
sym_PI(pars2)
```

##### Plot run 2 output
```{r run2_plots}
plot_run(run2)
```

##### Increase light to bleach tolerant symbiont
```{r}
env <- init_env(time=time, L=c(10,60,2), N=c(1e-7,1e-7,0), X=c(0,0,0))
run <- run_coral(time=time, env=env, pars=pars2)
plot_run(run)
```