### 2016-10-07 R/jCO2_analysis.Rmd
new model structure: jCw = CO2 input

---
title: "jCO2 analysis"
author: "Ross Cunning"
date: "October 5, 2016"
output: html_document
---

```{r load_functions, include=F, message=F}
sapply(c("../R/def_pars.R", "../R/init_env.R", "../R/plot_results.R", "../R/plot_steady_states.R"), source, .GlobalEnv)
```

```{r old_model_code, include=F}
run_coral <- function(time, env, pars, NPQ="inhibited", photoinhibition=T) {
  require(dplyr)
  # Define parallel complementary synthesizing unit function
  synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
  #synth <- function(x, y, m) (m^-2 + x^-2 + y^-2)^(-1/2)
  # Set initial values
  # ==================
  # Host
  H <- data_frame(
    time=time,
    H=pars$initH,
    jX=(pars$jXm * env$X[1] / (env$X[1] + pars$KX)),
    jCO2a=pars$jCO2a,
    jCO2p=pars$jCO2p,
    jCO2=jCO2a+jCO2p,
    jN=(pars$jNm * env$N[1] / (env$N[1] + pars$KN)),
    rhoN=jN,
    jCw=0,
    jHG=0.25,
    jHT=pars$jHT0,
    rNH=jHT * pars$nNH * pars$sigmaNH,
    rCH=jHT * pars$sigmaCH,
    dH.Hdt=0
  )
  H[2:nrow(H), 2:ncol(H)] <- NA

  # Symbiont
  S <- data_frame(
    time=time,
    S=pars$initS,
    jL=env$L[1] * pars$astar,
    jCP=max(0, synth(jL * pars$nLC, (H$jCO2p[1] + H$jCO2a[1])*H$H/S, pars$jCPm), na.rm=T),
    jeL=max(jL - jCP/pars$nLC, 0),
    jNPQ=pars$kNPQ/pars$nLC,
    jCO2w=H$jCO2*H$H/S - jCP,
    jSG=pars$jSGm/10,
    rhoC=jCP, 
    jNw=0,
    jST=pars$jST0,
    rNS=jST * pars$nNS * pars$sigmaNS,
    rCS=jST * pars$sigmaCS,
    cROS=1,
    dS.Sdt=1)
  S[2:nrow(S), 2:ncol(S)] <- NA

  # Run simulation by updating
  # ==========================
  for (t in 2:length(time)) {

    # Photosynthesis SU
    # =================
    # Light input flux
    S$jL[t] <- (1.256307 + 1.385969 * exp(-6.479055 * (S$S[t-1]/H$H[t-1]))) * env$L[t] * pars$astar
    # CO2 input flux
    H$jCO2p[t] <- pars$jCO2p # passive CO2 flux from the environment
    H$jCO2a[t] <- pars$jCO2a  # active delivery of CO2 from the environment (mediated by the host)
    H$jCO2[t] <- H$jCO2p[t] + H$jCO2a[t]  # Total CO2 input from the environment
    S$rCS[t] <- pars$jST0 * pars$sigmaCS  # metabolic CO2 recycled from symbiont biomass turnover
    H$rCH[t] <- H$jHT[t-1] * pars$sigmaCH  # metabolic CO2 recycled from host biomass turnover
    # Production flux (photosynthetic carbon fixation)
    if (photoinhibition==T) {
      S$jCP[t] <- synth(S$jL[t] * pars$nLC, (H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm) / S$cROS[t-1]
    } else if (photoinhibition==F) {
      S$jCP[t] <- synth(S$jL[t] * pars$nLC, (H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm)
    }

    # Rejection flux: CO2 (wasted to the environment)
    S$jCO2w[t] <- max((H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t] - S$jCP[t], 0)
    # Rejection flux: excess light energy not quenched by carbon fixation
    S$jeL[t] <- max(S$jL[t] - S$jCP[t]/pars$nLC, 0)
    # Amount of excess light energy quenched by NPQ
    #S$jNPQ[t] <- min(S$jeL[t], (S$jCP[t] * pars$kNPQ)/pars$nLC)
    if (NPQ=="inhibited") S$jNPQ[t] <- min(S$jeL[t], pars$kNPQ / S$cROS[t-1])
    if (NPQ=="inhibited2") S$jNPQ[t] <- (pars$kNPQ^(-3)+S$jeL[t]^(-3))^(-1/3) / S$cROS[t-1]
    if (NPQ=="fixed") S$jNPQ[t] <- min(S$jeL[t], pars$kNPQ)
    if (NPQ=="fixed2") S$jNPQ[t] <- (pars$kNPQ^(-3)+S$jeL[t]^(-3))^(-1/3)
    # Scaled ROS production due to excess excitation energy (=not quenched by carbon fixation AND NPQ)
    S$cROS[t] <- 1 + ((S$jeL[t] - S$jNPQ[t]) / pars$kROS)^pars$k

    # Symbiont biomass SU
    # ===================
    # Nitrogen input flux
    S$rNS[t] <- pars$jST0 * pars$nNS * pars$sigmaNS  # Recylced N from symbiont biomass turnover.
    H$rhoN[t-1] <- H$rhoN[t-1]  # Nitrogen shared from the host (defined below, so previous time step used)
    # Carbon input flux
    S$jCP[t] <- S$jCP[t]  # Production of fixed carbon from photosynthesis SU
    # Production flux (symbiont biomass formation)
    S$jSG[t] <- synth(S$jCP[t], (H$rhoN[t-1]*H$H[t-1]/S$S[t-1] + S$rNS[t])/pars$nNS, pars$jSGm)
    # Rejection flux: carbon (surplus carbon shared with the host)
    S$rhoC[t] <- max(S$jCP[t] - S$jSG[t], 0)
    # Rejection flux: nitrogen (surplus nitrogen wasted to the environment)
    S$jNw[t] <- max(H$rhoN[t-1]*H$H[t-1]/S$S[t-1] + S$rNS[t] - pars$nNS * S$jSG[t], 0)

    # Symbiont biomass loss (turnover)
    S$jST[t] <- pars$jST0 * (1 + pars$b * (S$cROS[t] - 1))

    # Host biomass SU
    # ===============
    # Food input flux (prey=both carbon and nitrogen)
    H$jX[t] <- (pars$jXm * env$X[t] / (env$X[t] + pars$KX))  # Prey uptake from the environment
    # Nitrogen input flux
    H$jN[t] <- (pars$jNm * env$N[t] / (env$N[t] + pars$KN))  # N uptake from the environment
    H$rNH[t] <- H$jHT[t-1] * pars$nNH * pars$sigmaNH  # Recycled N from host biomass turnover
    # Carbon input flux
    S$rhoC[t] <- S$rhoC[t]  # Carbon shared by the symbiont (defined above)
    # Production flux (host biomass formation)
    H$jHG[t] <- synth(S$rhoC[t]*S$S[t-1]/H$H[t-1] + H$jX[t], (H$jN[t] + pars$nNX*H$jX[t] + H$rNH[t]) / pars$nNH, pars$jHGm)
    # Rejection flux: nitrogen (surplus nitrogen shared with the symbiont)
    H$rhoN[t] <- max(H$jN[t] + pars$nNX * H$jX[t] + H$rNH[t] - pars$nNH * H$jHG[t], 0)
    # Rejection flux: carbon (surplus carbon lost to environment)
    H$jCw[t] <- max(H$jX[t] + S$rhoC[t]*S$S[t-1]/H$H[t-1] - H$jHG[t], 0)

    # Host biomass loss
    H$jHT[t] <- pars$jHT0

    # State equations
    # ===============
    # Specific growth rates (Cmol/Cmol/d)
    S$dS.Sdt[t] <- S$jSG[t] - S$jST[t]
    H$dH.Hdt[t] <- H$jHG[t] - H$jHT[t]
    # Biomass (Cmol)
    S$S[t] <- S$S[t-1] + S$dS.Sdt[t] * S$S[t-1] * (time[2] - time[1])
    H$H[t] <- H$H[t-1] + H$dH.Hdt[t] * H$H[t-1] * (time[2] - time[1])

  }

  # Return results
  # ==============
  return(list(time=time, env=env, pars=pars, H=H, S=S))
}
```

# Code for new model
The rejection flux of carbon from host biomass SU, $jCw$, is used as input to photosynthesis SU. A multiplier is included indicating the number of moles of CO~2~ the host can deliver to the photosynthesis SU using one mole of fixed carbon. Mechanistically, the host will use this carbon to generate ATP to energize a variety of CO~2~-concentrating mechanisms (CCM's) that enhance CO~2~ availability to photosynthesis.
```{r new_model, include=T}
run_coral_wC <- function(time, env, pars, NPQ="inhibited", jDOC=0) {
  require(dplyr)
  # Define parallel complementary synthesizing unit function
  synth <- function(x, y, m) 1/((1/m)+(1/x)+(1/y)-(1/(x+y)))
  #synth <- function(x, y, m) (m^-2 + x^-2 + y^-2)^(-1/2)
  # Set initial values
  # ==================
  # Host
  H <- data_frame(
    time=time,
    H=pars$initH,
    jX=(pars$jXm * env$X[1] / (env$X[1] + pars$KX)),
    jCO2a=pars$jCO2a,
    jCO2p=pars$jCO2p,
    jCO2=jCO2a+jCO2p,
    jN=(pars$jNm * env$N[1] / (env$N[1] + pars$KN)),
    rhoN=jN/2,
    jCw=1,
    jHG=0.25,
    jHT=pars$jHT0,
    rNH=jHT * pars$nNH * pars$sigmaNH,
    rCH=jHT * pars$sigmaCH,
    dH.Hdt=0
  )
  H[2:nrow(H), 2:ncol(H)] <- NA

  # Symbiont
  S <- data_frame(
    time=time,
    S=.2,
    jL=env$L[1] * pars$astar,
    jCP=max(0, synth(jL * pars$nLC, (H$jCw[1] + H$jCO2p[1] + H$jCO2a[1])*H$H/S, pars$jCPm), na.rm=T),
    jeL=max(jL - jCP/pars$nLC, 0),
    jNPQ=pars$kNPQ/pars$nLC,
    jCO2w=(H$jCw + H$jCO2)*H$H/S - jCP,
    jSG=pars$jSGm/10,
    rhoC=jCP/2, 
    jNw=0,
    jST=pars$jST0,
    rNS=jST * pars$nNS * pars$sigmaNS,
    rCS=jST * pars$sigmaCS,
    cROS=1,
    dS.Sdt=1)
  S[2:nrow(S), 2:ncol(S)] <- NA

  # Run simulation by updating
  # ==========================
  for (t in 2:length(time)) {

    # Photosynthesis SU
    # =================
    # Light input flux
    S$jL[t] <- (1.256307 + 1.385969 * exp(-6.479055 * (S$S[t-1]/H$H[t-1]))) * env$L[t] * pars$astar
    # CO2 input flux
    H$jCO2p[t] <- pars$jCO2p # passive CO2 flux from the environment
    H$jCO2a[t] <- pars$jCO2a  # active delivery of CO2 from the environment (mediated by the host)
    H$jCO2[t] <- H$jCO2p[t] + H$jCO2a[t]  # Total CO2 input from the environment
    S$rCS[t] <- pars$jST0 * pars$sigmaCS  # metabolic CO2 recycled from symbiont biomass turnover
    H$rCH[t] <- H$jHT[t-1] * pars$sigmaCH  # metabolic CO2 recycled from host biomass turnover
    H$jCw[t-1] <- H$jCw[t-1]  # carbon rejected from host biomass SU, given back as CO2 to photosynthesis
    # Production flux (photosynthetic carbon fixation)
    S$jCP[t] <- synth(S$jL[t] * pars$nLC, (5*H$jCw[t-1] + H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t], pars$jCPm) / S$cROS[t-1]

    # Rejection flux: CO2 (wasted to the environment)
    S$jCO2w[t] <- max((H$jCO2[t] + H$rCH[t])*H$H[t-1]/S$S[t-1] + S$rCS[t] - S$jCP[t], 0)
    # Rejection flux: excess light energy not quenched by carbon fixation
    S$jeL[t] <- max(S$jL[t] - S$jCP[t]/pars$nLC, 0)
    # Amount of excess light energy quenched by NPQ
    #S$jNPQ[t] <- min(S$jeL[t], (S$jCP[t] * pars$kNPQ)/pars$nLC)
    if (NPQ=="inhibited") S$jNPQ[t] <- min(S$jeL[t], pars$kNPQ / S$cROS[t-1])
    if (NPQ=="inhibited2") S$jNPQ[t] <- (pars$kNPQ^(-3)+S$jeL[t]^(-3))^(-1/3) / S$cROS[t-1]
    if (NPQ=="fixed") S$jNPQ[t] <- min(S$jeL[t], pars$kNPQ)
    if (NPQ=="fixed2") S$jNPQ[t] <- (pars$kNPQ^(-3)+S$jeL[t]^(-3))^(-1/3)
    # Scaled ROS production due to excess excitation energy (=not quenched by carbon fixation AND NPQ)
    S$cROS[t] <- 1 + ((S$jeL[t] - S$jNPQ[t]) / pars$kROS)^pars$k

    # Symbiont biomass SU
    # ===================
    # Nitrogen input flux
    S$rNS[t] <- pars$jST0 * pars$nNS * pars$sigmaNS  # Recylced N from symbiont biomass turnover.
    H$rhoN[t-1] <- H$rhoN[t-1]  # Nitrogen shared from the host (defined below, so previous time step used)
    # Carbon input flux
    S$jCP[t] <- S$jCP[t]  # Production of fixed carbon from photosynthesis SU
    # Production flux (symbiont biomass formation)
    S$jSG[t] <- synth(S$jCP[t], (H$rhoN[t-1]*H$H[t-1]/S$S[t-1] + S$rNS[t])/pars$nNS, pars$jSGm)
    # Rejection flux: carbon (surplus carbon shared with the host)
    S$rhoC[t] <- max(S$jCP[t] - S$jSG[t], 0)
    # Rejection flux: nitrogen (surplus nitrogen wasted to the environment)
    S$jNw[t] <- max(H$rhoN[t-1]*H$H[t-1]/S$S[t-1] + S$rNS[t] - pars$nNS * S$jSG[t], 0)

    # Symbiont biomass loss (turnover)
    S$jST[t] <- pars$jST0 * (1 + pars$b * (S$cROS[t] - 1))

    # Host biomass SU
    # ===============
    # Food input flux (prey=both carbon and nitrogen)
    H$jX[t] <- (pars$jXm * env$X[t] / (env$X[t] + pars$KX))  # Prey uptake from the environment
    # Nitrogen input flux
    H$jN[t] <- (pars$jNm * env$N[t] / (env$N[t] + pars$KN))  # N uptake from the environment
    H$rNH[t] <- H$jHT[t-1] * pars$nNH * pars$sigmaNH  # Recycled N from host biomass turnover
    # Carbon input flux
    S$rhoC[t] <- S$rhoC[t]  # Carbon shared by the symbiont (defined above)
    # Add dissolved organic carbon uptake
    # Production flux (host biomass formation)
    H$jHG[t] <- synth(S$rhoC[t]*S$S[t-1]/H$H[t-1] + H$jX[t] + jDOC, (H$jN[t] + pars$nNX*H$jX[t] + H$rNH[t]) / pars$nNH, pars$jHGm)
    # Rejection flux: nitrogen (surplus nitrogen shared with the symbiont)
    H$rhoN[t] <- max(H$jN[t] + pars$nNX * H$jX[t] + H$rNH[t] - pars$nNH * H$jHG[t], 0)
    # Rejection flux: carbon -- given back to symbiont as CO2 input to photosynthesis
    H$jCw[t] <- max(H$jX[t] + S$rhoC[t]*S$S[t-1]/H$H[t-1] - H$jHG[t], 0)

    # Host biomass loss
    H$jHT[t] <- pars$jHT0

    # State equations
    # ===============
    # Specific growth rates (Cmol/Cmol/d)
    S$dS.Sdt[t] <- S$jSG[t] - S$jST[t]
    H$dH.Hdt[t] <- H$jHG[t] - H$jHT[t]
    # Biomass (Cmol)
    S$S[t] <- S$S[t-1] + S$dS.Sdt[t] * S$S[t-1] * (time[2] - time[1])
    H$H[t] <- H$H[t-1] + H$dH.Hdt[t] * H$H[t-1] * (time[2] - time[1])

  }

  # Return results
  # ==============
  return(list(time=time, env=env, pars=pars, H=H, S=S))
}
```

# Comparison of new and old model behavior
## Varying nitrogen
### Old model
```{r vary_DIN_old, cache=T, message=F, warning=F}
# Set up simulation
time <- seq(1,1230,1)
defpars <- def_pars()
defpars <- replace(defpars, "jHGm", 1)
env <- init_env(time=time, L=c(15,15,2), N=c(1e-7,1e-6,2), X=c(0,0,0))

# Run old model (uses jCO2a and jCO2p to deliver CO2 to photosynthesis SU)
run <- run_coral(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jHGm"), 
                              c(60,       60,    1,   5,      0.9,      0.9,      0.32,     1)),
                 NPQ="fixed2")
plot_run(run)
```

### New model
```{r vary_DIN_new, cache=T, message=F, warning=F}
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm"), 
                              c(60,       60,    1,   5,      0.9,      0.9,       0,       0,      1)),
                 NPQ="fixed2")
plot_run_wC(run)
```

## Varying light
### Old model
```{r vary_L_old, cache=T, message=F, warning=F}
# Set up simulation
time <- seq(1,1230,1)
env <- init_env(time=time, L=c(10,25,2), N=c(1e-7,1e-7,2), X=c(0,0,0))

# Run old model (uses jCO2a and jCO2p to deliver CO2 to photosynthesis SU)
run <- run_coral(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(60,       60,    1,   5,      0.9,     0.9,      0.32)),
                 NPQ="fixed2")
plot_run(run)
```

### New model
```{r vary_L_new, cache=T, message=F, warning=F}
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)
```

## Light stress
### old model: Increasing Light in a non-feeding host
```{r incr_L_old_nofeeding, cache=T, message=F, warning=F}
# Set up simulation
env <- init_env(time=time, L=c(2,60,0), N=c(1e-7,1e-7,0), X=c(0,0,0))
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0.32)),
                 NPQ="fixed2")
plot_run(run)
```

### new model: Increasing Light in a non-feeding host
```{r incr_L_nofeeding, cache=T, message=F, warning=F}
# Set up simulation
env <- init_env(time=time, L=c(2,60,0), N=c(1e-7,1e-7,0), X=c(0,0,0))
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)
```

### new model: Increasing Light in a feeding host
```{r incr_L, cache=T, message=F, warning=F}
# Set up simulation
env <- init_env(time=time, L=c(2,60,0), N=c(1e-7,1e-7,0), X=c(5e-6,5e-6,0))
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)
```

#### possibility of DOC uptake?
```{r incr_L_DOC, cache=T, message=F, warning=F, include=F, eval=F}
# Set up simulation
env <- init_env(time=time, L=c(2,60,0), N=c(1e-7,1e-7,0), X=c(0,0,0))
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env, jDOC=0.005,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)
```

### Light stress with old model
```{r Lstress_old, cache=T, message=F, warning=F}
# Set up simulation
time <- seq(1,1230,1)
env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,2), X=c(0,0,0))

# Run old model (uses jCO2a and jCO2p to deliver CO2 to photosynthesis SU)
run <- run_coral(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(60,       60,    1,   5,      0.9,     0.9,      0.32)),
                 NPQ="fixed2")
plot_run(run)
```

### Light stress with new model
```{r Lstress_new, cache=T, message=F, warning=F}
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)
```

#### Examine sources of CO~2~ for photosynthesis in new model
```{r CO2_sources, cache=T, message=F, warning=F, fig.width=3.5, fig.height=6}
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw and metabolic CO2)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
par(mfrow=c(4,1), mar=c(2,2,1,1))
# Plot light
plot_L(run)
# Plot symbiont to host ratio
plot_sh(run)
# Plot symbiont-specific CO2 delivery sources
with(run, {
  totalCO2 <- (H$jCw+H$rCH)*H$H/S$S + S$rCS
  plot(time, totalCO2, type="l", col="gray")
  lines(time, H$jCw*H$H/S$S, lty=1, col="black")
  lines(time, H$rCH*H$H/S$S, lty=2, col="brown")
  lines(time, S$rCS, lty=2, col="green")
})
# Plot proportions of each CO2 source
with(run, {
  totalCO2 <- (H$jCw+H$rCH)*H$H/S$S + S$rCS
  plot(time, H$jCw*H$H/S$S/totalCO2, type="l", ylim=c(0,1))
  lines(time, H$rCH*H$H/S$S/totalCO2, lty=2, col="brown")
  lines(time, S$rCS/totalCO2, lty=2, col="green")
})
```

### Light stress with new model, +feeding
```{r Lstress_new_feeding, cache=T, message=F, warning=F}
# Set up simulation
env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,2), X=c(1e-6,1e-6,0))
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)
```

### Light stress with new model, +feeding, reduced DIN
```{r Lstress_new_feeding_lowerDIN, cache=T, message=F, warning=F}
# Set up simulation
env <- init_env(time=time, L=c(10,40,2), N=c(0.5e-7,0.5e-7,2), X=c(1e-6,1e-6,0))
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)
```

### Light stress with new model, +tolerant symbiont
```{r Lstress_new_tolerant, cache=T, message=F, warning=F}
# Set up simulation
env <- init_env(time=time, L=c(10,40,2), N=c(1e-7,1e-7,2), X=c(1e-6,1e-6,0))
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       120,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)
```

## Increasing DIN
### Old model
```{r incr_DIN_old, cache=T, message=F, warning=F}
# Set up simulation
env <- init_env(time=time, L=c(30,30,0), N=c(1e-8,4e-6,0), X=c(0,0,0))
# Run old model
run <- run_coral(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0.32)),
                 NPQ="fixed2")
plot_run(run)
```

### New model
```{r incr_DIN_new, cache=T, message=F, warning=F}
# Set up simulation
env <- init_env(time=time, L=c(30,30,0), N=c(1e-8,4e-6,0), X=c(0,0,0))
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0)),
                 NPQ="fixed2")
plot_run_wC(run)
```

# Steady state behavior of new model
```{r new_steadystates_code, message=F, warning=F, include=F}
run_steady_states <- function(time, pars, at, outfile=NULL, runtime=F, food=0, jDOC=0) {
  #
  run_coral_wC <- get("run_coral_wC")
  # Load libraries
  require(foreach)
  require(parallel)
  require(doParallel)
  require(reshape2)
  require(dplyr)

  # Load functions
  source("../R/init_env.R", local=TRUE)
  #source("R/run_coral.R", local=TRUE)

  # Set up cluster for parallel processing
  cl <- makeCluster(detectCores())  # Initiate cluster
  registerDoParallel(cl)

  # Run simulations in parallel for each combination of "at" values
  start <- proc.time()  # Start timer...
  # Calculate steady states...
  steady_states <- foreach(i=1:nrow(at), .combine='rbind', .packages='dplyr') %dopar% {
    L <- at[i,]$L; N <- at[i,]$N
    env <- init_env(time=time, L=c(L,L,0), N=c(N,N,0), X=c(food,food,0))
    run <- run_coral_wC(time=time, env=env, pars=pars, jDOC=jDOC)
    ss <- lapply(run[c("H", "S")], function(x) x[length(time), ])
    gr <- ss$H$dH.Hdt
    sh <- ss$S$S/ss$H$H
    hl <- with(ss, log(  pmin(S$rhoC*S$S/H$H + H$jX, pars$jHGm) / pmin((H$jN + pars$nNX*H$jX + H$rNH) / pars$nNH, pars$jHGm)  ))
    sl <- with(ss, log(  pmin(S$jCP, pars$jSGm)   /   pmin((H$rhoN*H$H/S$S + S$rNS)/pars$nNS, pars$jSGm)  ))
    ee <- with(ss, max(0, S$jL - (S$jCP/run$pars$nLC + run$pars$kNPQ)))
    pl <- with(ss, log(   pmin((H$jCw + H$jCO2 + H$rCH)*H$H/S$S + S$rCS, pars$jCPm)   /   pmin(S$jL * pars$nLC, pars$jCPm)    ))
    data.frame(gr=gr, sh=sh, hl=hl, sl=sl, ee=ee, pl=pl)
  }
  stopCluster(cl)  # Stop cluster
  if (runtime==T) print(proc.time() - start)  # Print time elapsed if runtime==T

  # Collect results in ss
  ss <- cbind(at, steady_states)

  # If output filename give, save ss as Rdata
  if (!is.null(outfile)) save(ss, file=outfile)

  # Return steady state values
  return(ss)
}
```

## No feeding
```{r run_new_steadystates, cache=T, message=F, warning=F}
# Set time and parameters 
time <- seq(1,100,1) 
pars <- replace(defpars, c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                          c(60,       60,    1,   5,      0.9,     0.9,       0,       0))
# Set values of L and N at which to get steady state values 
at <- expand.grid(L=seq(from=0, to=60, length.out=41),
                  N=seq(from=0, to=4e-6, length.out=41))

# Run to steady state and plot results
ssdat <- run_steady_states(time=time, pars=pars, at=at, food=0)
plot_steady_states(ssdat)
```

## + feeding
```{r run_new_steadystates_feeding, cache=T, message=F, warning=F}
# Set time and parameters 
time <- seq(1,100,1)
pars <- replace(defpars, c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                          c(60,       60,    1,   5,      0.9,     0.9,       0,       0))
# Set values of L and N at which to get steady state values 
at <- expand.grid(L=seq(from=0, to=60, length.out=41),
                  N=seq(from=0, to=4e-6, length.out=41))

# Run to steady state and plot results
ssdat <- run_steady_states(time=time, pars=pars, at=at, food=5e-6)
plot_steady_states(ssdat)
```

## + feeding + tolerant symbiont
```{r run_new_steadystates_feeding_tolerant, cache=T, message=F, warning=F}
# Set time and parameters 
time <- seq(1,100,1)
pars <- replace(defpars, c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p"), 
                          c(60,       200,    1,   5,      0.9,     0.9,       0,       0))
# Set values of L and N at which to get steady state values 
at <- expand.grid(L=seq(from=0, to=60, length.out=41),
                  N=seq(from=0, to=4e-6, length.out=41))

# Run to steady state and plot results
ssdat <- run_steady_states(time=time, pars=pars, at=at, food=5e-6)
plot_steady_states(ssdat)
```

# Stability at high nitrogen...?
## When system starts at 2 µM DIN and increases to 4 µM DIN, outcome is negative growth.
```{r incrN1, cache=T, message=F, warning=F}
# Set up simulation
env <- init_env(time=time, L=c(10,10,0), N=c(2e-6,4e-6,0), X=c(0,0,0))
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 1)),
                 NPQ="fixed2")
plot_run_wC(run)
```

## When system starts at 1 µM DIN and increases to 4 µM DIN, outcome is positive growth.
```{r incrN2, cache=T, message=F, warning=F}
# Set up simulation
env <- init_env(time=time, L=c(10,10,0), N=c(1e-6,4e-6,0), X=c(0,0,0))
# Run new model (sets jCO2a and jCO2p to zero; CO2 is delivered by jCw)
run <- run_coral_wC(time=time, env=env,
                 pars=replace(defpars, 
                              c("kNPQ", "kROS", "k", "b", "sigmaCS", "sigmaCH", "jCO2a", "jCO2p", "jHGm"), 
                              c(60,       60,    1,   5,      0.9,         0.9, 0, 0, 1)),
                 NPQ="fixed2")
plot_run_wC(run)
```