### 2017-01-08 JTB_ms/JTB_ms.Rmd
Finish figure 8 and overshoot discussion

---
title: A simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses and coral bleaching as an alternate stable state
author:
  - name: Ross Cunning
    email: ross.cunning@gmail.com
    affiliation: University of Hawaii
    footnote: Corresponding Author
  - name: Erik B. Muller
    affiliation: University of California
  - name: Ruth D. Gates
    affiliation: University of Hawaii
  - name: Roger M. Nisbet
    affiliation: University of California
address:
  - code: University of Hawaii
    address: Hawaii Institute of Marine Biology, Kaneohe, HI 96744, USA
  - code: University of California
    address: Department of Ecology, Evolution, and Marine Biology, Santa Barbara, CA 93106, USA
abstract: |
  Coral reef ecosystems owe their ecological success--and vulnerability to climate change--to the symbiotic metabolism of corals and *Symbiodinium* spp. The urgency to understand and predict the stability and breakdown of these symbioses (i.e., coral 'bleaching') demands the development and application of theoretical tools; however, the complexity of some approaches may limit their potential application and utility. Here, we develop a simplified dynamic bioenergetic model of coral-*Symbiodinium* symbioses that demonstrates realistic steady-state patterns in coral growth and symbiont abundance across gradients of light, nutrients, and feeding. Furthermore, through a mechanistic treatment of photo-oxidative stress, the model displays dynamic behaviors of bleaching and recovery, and suggests that these phenomena represent transitions between alternate stable states. A suite of complex responses to multiple, interacting environmental factors reproduced by the model suggests it unifyingly captures many important attributes of the system; meanwhile, its modular framework and open source R code are designed to facilitate expansion and problem-specific applications. We see significant potential applications for this modeling framework in generating testable hypotheses and predicting integrated, mechanistic responses of corals to environmental change.
bibliography: library.bib
output: 
  rticles::elsevier_article:
    fig_caption: yes
---
```{r setup, include=FALSE}
library(reshape2); library(dplyr)
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r load, include=FALSE}
# Source metrics reported in text
source("R/reported_metrics.R")
# Source default parameter values
source("R/def_pars.R")
defpars <- def_pars()
```

# Introduction

The nutritional exchange between corals and *Symbiodinium* directly underlies the capacity of corals to build coral reef ecosystems, worth trillions of US Dollars annually [@Costanza:2014ex]. However, the complex symbiotic metabolism of corals is vulnerable to disruption by numerous anthropogenic environmental perturbations, jeopardizing their future persistence. In order to understand and predict responses of corals to complex changes in their environment, a mechanistic understanding of how multiple interacting factors drive the individual and emergent physiology of both symbiotic partners is necessary. Such a task is well suited for theoretical modeling frameworks such as Dynamic Energy Budget (DEB) theory [@Kooijman:2010vd], although the complexity of such theory makes these efforts inaccessible to many biologists [@Jager:2013bj]. In order to bridge this gap, we present here a simplified dynamic bioenergetic model for coral-*Symbiodinium* symbioses that aims to mechanistically integrate the impacts of complex environmental change on the physiological performance of reef corals, including responses to environmental stress.

In reef coral symbioses, intracellular *Symbiodinium* translocate photosynthetically fixed carbon to support coral metabolism, while the animal host provides access to inorganic nutrients and carbon dioxide [@Muscatine:1977p4220]. Previous application of DEB theory to this syntrophic system [@Muller:2009io] demonstrated a stable symbiotic relationship and qualitatively realistic growth and biomass ratios across gradients of ambient irradiance, nutrients, and food. This model assumed that 1) *Symbiodinium* has priority access to fixed carbon through photosynthesis, 2) the coral animal has priority access to dissolved nitrogen through contact with seawater, and 3) each partner shares with the other only what it cannot use for its own growth. In its simplest form, this principle of sharing the surplus is sufficient to describe the dynamics of diverse syntrophic organs and organisms (e.g., trees, duckweeds, corals), suggesting the mechanism is mathematically and evolutionarily robust (Nisbet et al., in prep.).

While the formal DEB model of @Muller:2009io represents the most significant theoretical contribution in coral symbiosis research to date, we aim to strengthen the role of theory and broaden its potential application in three primary ways:

1. _Develop a module of photooxidative stress._ Of primary interest to coral biologists and ecologists is symbiosis dysfunction under environmental stress, resulting in coral “bleaching”--the loss of algal symbionts from the association [@Jokiel:1977p7353]. Photooxidative stress in *Symbiodinium* is considered a primary trigger of bleaching in response to high temperature and/or light [@Weis:2008p944], and prolonged or severe bleaching can result in mortality, though corals sometimes recover their symbionts. Bleaching susceptibility, severity, and recovery may be influenced by interacting factors such as heterotrophy and nutrient availability [@Wooldridge:2014hc]. To simulate these events and interactions, we develop a photooxidative stress module linking overreduction of the photosynthetic light reactions to downstream impacts of photoinhibition and photodamage.

2. _Reduce theoretical and mathematical complexity._ Following the logic of @Jager:2013bj, we exclude certain features of formal DEB theory in order to capture behaviors of interest with the simplest possible formulation. Here, we present a model without reserves, maturity, or reproduction [see @Kooijman:2010vd]. This formulation restricts the model's scope to the bioenergetics of adult corals (i.e., reproduction, larval stages, and metamorphosis are not considered), but greatly reduces theoretical complexity and parameter numbers, which is advantageous given the relative paucity of data for corals. However, our primary motivation for reducing complexity was to increase accessibility and applicability for biologists and ecologists without requiring significant expertise in DEB theory.

3. _Provide well-documented, open-access code._ In order to facilitate the continued development and application of theoretical modeling tools for coral symbioses, we provide open access to the model in the form of detailed, commented code written in the R language [@RALanguageandEn:2014wf]. With an accessible and modular framework, we envision this model as a resource for futher development by the scientific community to include additional complexity and problem-specific components. We chose the R language, which is freely available and in common use by biologists and ecologists, to reach the widest possible audience with this work.

With these as our primary motivations, we describe a simplified approach to dynamic bioenergetic modeling of coral-*Symbiodinium* symbioses that dynamically integrates the influence of external irradiance, nutrients, and prey availability on coral growth and symbiosis dynamics (i.e., symbiont:host biomass ratios), allowing for the possibility of coral bleaching in the event of photooxidative stress. In the following sections, we describe and provide rationale for the model structure, and demonstrate a range of steady state and dynamic behaviors that are consistent with observed phenomena.

# Model description
In this dynamical system, both partners acquire and use carbon and nitrogen to construct biomass: the symbiont fixes carbon through photosynthesis, and receives nitrogen shared by the host, while the host acquires nitrogen from the environment and receives carbon shared by the symbiont. A graphical representation of the model is presented in Fig. 1, and each model flux and parameter is defined in Tables 1 and 2, respectively. We use C-moles as the unit of biomass for consistency with the rigorous mass balance of DEB theory: 1 C-mole is equivalent to the amount of biomass containing 1 mole of carbon atoms. Host biomass ($H$), symbiont biomass ($S$), and prey biomass ($X$) have fixed, but different, molar N:C ratios (Table 2). Biomass is produced from carbon and nitrogen by synthesizing units (SU), which are mathematical specifications of the formation of a product from two substrates; we use the parallel complementary formulation of @Kooijman:2010vd to specify these fluxes. The two state variables of this system are symbiont biomass and coral biomass; because resources are acquired proportionally to surface area, and surface area is proportional to volume (i.e., corals are "V1-morphs" in DEB terminology [@Kooijman:2010vd]), biomass increases exponentially during growth (indeed, corals grow exponentially [@Bak:1976bv]).

Environmental stress is implemented in the form of photooxidative stress, which is thought to be the primary trigger of coral bleaching [@Lesser:1997p3587; @Weis:2008p944; @Wooldridge:2009p7807]. To simulate these events, we model the absorption and quenching of light energy by photochemistry and non-photochemical quenching, and the phenomenological responses that occur when these capacities are overwhelmed (i.e., photoinhibition, photodamage, and symbiont loss). While bleaching in response to high light alone has been observed experimentally [@Schutter:2011ie; @Downs:2013kc], mass coral bleaching events occur in response to high temperature [@HoeghGuldberg:1999p4222]; thus, it is important to justify our consideration of light as the primary stressor. In reality, light and temperature interact synergistically [@Coles:1978p1124; @Jones:1998p3572], and in fact, any stressor that disrupts the quenching of light energy may lead to bleaching [@Wooldridge:2010p7809; @Baker:2015kp]. This is because the proximate cause of photo-oxidative stress is excess excitation energy, but the upstream events that lead to this situation may be diverse: indeed, elevated temperature may inhibit Rubisco functioning [@Jones:1998p3572] and the repair of the D1 protein in photosystem II [@Warner:1999p4239], which reduces the capacity of photochemical quenching and leads to an excess of light energy. In this way, elevated temperature serves to reduce the threshold above which light stresses the system [@HoeghGuldberg:1999p4222]; importantly, light is still the proximate stressor. Therefore, we omitted temperature from the model to maintain a desired level of simplicity, while still allowing photooxidative stress and bleaching to be simulated with biological realism in response to light.

Below we describe the formulation of each flux in the model.

Table 1. Model fluxes (mass-specific)

| Symbol  | Description                            | Units               | Eq. no. |
| ------- | -------------------------------------- | ------------------- | ------- |
| $j_X$     | Prey uptake rate                       | molX CmolH^-1^ d^-1^ | 3 |
| $j_N$     | Nitrogen uptake rate                   | molN CmolH^-1^ d^-1^ | 4 |
| $j_{HG}$  | Host biomass formation rate            | CmolH CmolH^-1^ d^-1^| 5 |
| $r_{NH}$  | Recycled nitrogen from host turnover   | molN CmolH^-1^ d^-1^ | 6 |
| $\rho_N$  | Nitrogen shared with the symbiont      | molN CmolH^-1^ d^-1^ | 7 |
| $j_{eC}$  | Excess carbon used to activate host CCMs| molC CmolH^-1^ d^-1^ | 8 |
| $j_{CO_2}$| CO~2~ input to photosynthesis          | molCO~2~ CmolH^-1^ d^-1^| 9 |
| $j_L$     | Light absorption rate                  | molph CmolS^-1^ d^-1^| 11 |
| $r_{CH}$  | Recycled CO~2~ from host turnover      | molCO~2~ CmolH^-1^ d^-1^| 12 |
| $r_{CS}$  | Recycled CO~2~ from symbiont turnover  | molCO~2~ CmolS^-1^ d^-1^| 13 |
| $j_{CP}$  | Photosynthesis rate                    | molC CmolS^-1^ d^-1^ | 14 |
| $j_{eL}$  | Light energy in excess of photochemistry| molph CmolS^-1^ d^-1^| 15 |
| $j_{NPQ}$ | Total capacity of NPQ                  | molph CmolS^-1^ d^-1^ | 16 |
| $c_{ROS}$ | ROS production proportional to baseline | -- | 17 |
| $r_{NS}$  | Recycled nitrogen from symbiont turnover| molN CmolS^-1^ d^-1^| 18 |
| $j_{SG}$  | Symbiont biomass formation rate        | CmolS CmolS^-1^ d^-1^| 19 |
| $\rho_C$  | Fixed carbon shared with host          | molC CmolS^-1^ d^-1^ | 20 |
| $j_{ST}$  | Symbiont turnover rate                 | CmolS CmolS^-1^ d^-1^| 21 |

Table 2. Model parameters. Derivations of each parameter value along with supporting references are provided in the Supplementary Information.

| Symbol  | Description                            | Units               | Value   |
| ------- | -------------------------------------- | ------------------- | ------- |
| $n_{NH}$ | N:C molar ratio in host biomass | -- | $`r defpars$nNH`$ |
| $n_{NS}$ | N:C molar ratio in symbiont biomass | -- | $`r defpars$nNS`$ |
| $n_{NX}$ | N:C molar ratio in prey biomass | -- | $`r defpars$nNX`$ |
| $j_{HT}^0$ | Specific turnover rate of host biomass | CmolH CmolH^-1^ d^-1^ | $`r defpars$jHT0`$ |
| $j_{ST}^0$ | Specific turnover rate of symbiont biomass | CmolS CmolS^-1^ d^-1^ | $`r defpars$jST0`$ |
| $\sigma_{NH}$ | Proportion host N-turnover recycled to host | -- | $`r defpars$sigmaNH`$ |
| $\sigma_{CH}$ | Proportion host C-turnover recycled to photosynthesis | -- | $`r defpars$sigmaCH`$ |
| $\sigma_{NS}$ | Proportion symbiont N-turnover recycled to symbiont | -- | $`r defpars$sigmaNS`$ |
| $\sigma_{CS}$ | Proportion symbiont C-turnover recycled to photosynthesis | -- | $`r defpars$sigmaCS`$ |
| $j_{Xm}$ | Maximum specific feeding rate of host | molX CmolH^-1^ d^-1^ | $`r defpars$jXm`$ |
| $K_X$ | Half-saturation constant for prey uptake by host | molX L^-1^ | $`r defpars$KX`$ |
| $j_{Nm}$ | Maximum specific DIN uptake rate by host | molN CmolH^-1^ d^-1^ | $`r defpars$jNm`$ |
| $K_N$ | Half-saturation constant for DIN uptake by host | molN L^-1^ | $`r defpars$KN`$ |
| $k_{CO_2}$ | Efficacy of CO~2~ delivery by host CCMs | molCO~2~ molC^-1^ | $`r defpars$kCO2`$ |
| $j_{HGm}$ | Maximum specific growth rate of host | CmolH CmolH^-1^ d^-1^ | $`r defpars$jHGm`$ |
| $y_{CL}$ | Quantum yield of photosynthesis | molC mol ph^-1^ | $`r defpars$yCL`$ |
| $\bar{a}^*$ | Effective light-absorbing cross-section of symbiont | m^2^ CmolS^-1^ | $`r defpars$astar`$ |
| $k_{NPQ}$ | NPQ capacity of symbiont | mol ph CmolS^-1^ d^-1^ | $`r defpars$kNPQ`$ |
| $k_{ROS}$ | Excess photon energy that doubles ROS production, relative to baseline levels | mol ph CmolS^-1^ d^-1^ | $`r defpars$kROS`$ |
| $j_{CPm}$ | Maximum specific photosynthesis rate of symbiont | molC CmolS^-1^ d^-1^ | $`r defpars$jCPm`$ |
| $j_{SGm}$ | Maximum specific growth rate of symbiont | CmolS CmolS^-1^ d^-1^ | $`r defpars$jSGm`$ |
| $b$ | Scaling parameter for strength of bleaching response | -- | $`r defpars$b`$ |

![Graphical representation of coral-algal symbiosis model. Light, CO~2~, prey, and DIN are acquired from the external environment proportional to the biomass of the partner indicated by the black box for uptake. Mass fluxes (see Table 1 for definitions) are represented by $j$'s with subscripts indicating the type of mass, and in some cases the process (e.g., $j_{CP}$ is the flux of carbon produced by photosynthesis), and $\rho$'s indicate fluxes that are shared by one partner with the other. Parallel complementary synthesizing units (SUs) are represented by large circles, and rejection fluxes from these SUs are indicated by dashed lines. $c_{ROS}$ is a proportional rate that impacts other model fluxes by inhibition or acceleration; likewise, $j_{eC}$ accelerates the rate of $j_{CO_2}$. Recycled mass fluxes from biomass turnover are not shown for clarity (but see Table 1 for definitions).](../img/Fig1.png)

## State equations
The balance equations for symbiont ($S$) and host ($H$) biomass are expressed as:

\begin{equation} {dS \over Sdt} = j_{SG} - j_{ST} \end{equation}

\begin{equation} {dH \over Hdt} = j_{HG} - j_{HT}^0 \end{equation}

These are expressed per unit of symbiont and host biomass, respectively. The specific biomass growth and turnover rates that define these balance equations are produced by combinations of the individual model fluxes (see Table 1 for definitions and units), which are each expressed as mass-specific rates (e.g., per C-mole of symbiont or host biomass per day). When necessary, conversions between symbiont-mass-specific and host-mass-specific rates are accomplished by multiplying or dividing by the symbiont:host biomass ratio.

## Coral animal fluxes

The coral animal acquires both carbon and nitrogen from feeding on prey from the environment. Prey acquisition is specified by Michaelis-Menten kinetics (i.e., a Holling type II function) with a maximum feeding rate $j_{Xm}$ and half-saturation constant $K_X$:

\begin{equation} j_X = {{j_{Xm} \cdot X} \over {X + K_X}} \end{equation}

Additionally, the coral animal acquires nitrogen dissolved in the surrounding seawater, which is assumed to represent ammonium, the primary form utilized by corals [@Yellowlees:2008p331]. This gives the host (rather than the symbiont) priority in nitrogen utilization; this capacity is supported by experimental evidence [@Wang:1998p128] and is consistent with the physical arrangment of the partners, where the host is in direct contact with the external environment. The uptake of nitrogen from the environment is thus specified by Michaelis-Menten kinetics using a maximum uptake rate $j_{Nm}$ and half-saturation constant $K_N$:

\begin{equation} j_N = {{j_{Nm} \cdot N} \over {N + K_N}} \end{equation}

Coral biomass formation is then specified by a parallel complementary SU that combines carbon and nitrogen, according to:

\begin{equation} j_{HG} = \bigg({1 \over j_{HGm}} + {1 \over {\rho_C{S \over H} + j_X}} + {1 \over {(j_N + n_{NX}j_X + r_{NH}) / n_{NH}}} - {1 \over {{\rho_C{S \over H} + j_X} + (j_N + n_{NX}j_X + r_{NH}) / n_{NH}}}\bigg)^{-1} \end{equation}

where $\rho_C$ is fixed carbon shared by the symbiont (see Eq. 20), and $r_{NH}$ is recycled nitrogen liberated by host biomass turnover:

\begin{equation} r_{NH}=\sigma_{NH}n_{NH}j_{HT}^0 \end{equation}

The amount of nitrogen input to the coral biomass SU in excess of what is actually consumed in biomass formation (i.e., surplus nitrogen, or the rejection flux[^1] of the SU) is then made available to the symbiont:

\begin{equation} \rho_N = (j_N + n_{NX}j_X + r_{NH} - n_{NH}j_{HG})_+ \end{equation}

[^1]: Rejection fluxes must always be positive, and hence are specified with the notation $(x)_+$, which means $max(x, 0)$.

Due to the inherent inefficiency of the parallel complementary SU formulation, there is always some nitrogen shared with the symbiont even when coral biomass formation is strongly nitrogen-limited. Likewise, there is always a non-zero rejection flux of excess carbon from the coral biomass SU. Likewise, the carbon rejected from this SU reflects the amount of excess fixed carbon available to the host that is not used in biomass formation:

\begin{equation} j_{eC} = (jX + \rho_C{S \over H} - j_{HG})_+ \end{equation}

This flux, $j_{eC}$, is assumed to be available to the host as a respiratory substrate to support various energetically-demanding processes; of particular importance is the host's active carbon concentrating mechanisms (CCMs) that supply CO~2~ for symbiont photosynthesis [@Wooldridge:2013tj; @Hopkinson:2015ev]. We therefore specify $j_{CO_2}$ as the host-mediated delivery of CO~2~ to photosynthesis that encompasses potentially diverse carbon-concentrating mechanisms (CCMs), including active transport of bicarbonate, carbonic anhydrase-catalyzed conversion of bicarbonate to CO~2~ to promote diffusion toward the symbiont [@Tansik:2015cm], and acidification of the symbiosome to increase localized CO~2~ concentrations around the symbiont [@Barott:2014gx]. Since these active CCMs require energetic input by the host, we define $j_{CO_2}$ as proportional to $j_{eC}$, assuming that some of this carbon is respired to energize the CCMs. This formulation means that the symbiont indirectly ensures its own CO~2~ supply by providing fixed carbon (=energy) to the host [@Wooldridge:2013tj], which establishes a positive feedback in the model that leads to the existence of multiple stable states, which we discuss below in the context of coral bleaching. The parameter $k_{CO_2}$ scales the efficacy of host CCMs, which enables the comparison of different rates of CO~2~ delivery that may characterize different coral species [@Wooldridge:2014di]. The active input of CO~2~ to the photosynthesis SU is therefore specified as:

\begin{equation} j_{CO_2} = k_{CO_2}j_{eC} \end{equation}

## *Symbiodinium* fluxes

The symbiont produces fixed carbon through photosynthesis, a process represented here by a single SU with two substrates: light (photons) and inorganic carbon (CO~2~). The amount of light absorbed by the symbiont depends on the scalar irradiance at the site of light absorption, which is modified substantially relative to external downwelling irradiance owing to multiple scattering by the coral skeleton and self-shading by surrounding symbionts [@Enriquez:2005p142; @Marcelino:2013hz]. We used data from @Marcelino:2013hz to empirically derive an amplification factor, $A$, indicating the ratio of internal scalar irradiance to external downwelling irradiance as a function of symbiont density (S:H biomass), which is specified as:

\begin{equation} A = 1.26 + 1.39 \cdot \exp(-6.48 \cdot {S \over H}) \end{equation}

This amplification factor is then multiplied by the external downwelling irradiance $L$ and the effective light-absorbing surface area of symbiont biomass $\bar{a}^*$ to specify the total light absorption:

\begin{equation} j_L =  A \cdot L \cdot \bar{a}^* \end{equation}

CO~2~ arrives at the photosynthesis SU from multiple sources: in addition to the CO~2~ actively supplied by the host through its CCMs ($j_{CO_2}$; Eq. 9), a fixed proportion of metabolic CO~2~ generated by host biomass turnover is passively available to the photosynthesis SU, according to:

\begin{equation} r_{CH}=\sigma_{CH}j_{HT}^0 \end{equation}

along with a fixed proportion of CO~2~ generated by symbiont biomass turnover:

\begin{equation} r_{CS}=\sigma_{CS}j_{ST}^0 \end{equation}

Fixed carbon is then produced by the photosynthesis SU according to:

\begin{equation} j_{CP} = \bigg({1 \over j_{CPm}} + {1 \over {y_{CL} j_L }} + {1 \over {(j_{CO_2} + r_{CH}){H \over S} + r_{CS}}} - {1 \over {y_{CL} j_L + (j_{CO_2} + r_{CH}){H \over S} + r_{CS}}}\bigg)^{-1} \cdot c_{ROS}^{-1} \end{equation}

where $j_{CPm}$ is the maximum specific rate of photosynthesis, and $c_{ROS}$ is the relative rate of reactive oxygen species production (see Eq. 17). Dividing the photosynthetic rate by $c_{ROS}$ causes a decline in response to photooxidative stress at high light levels, and the emergent outcome of this SU formulation demonstrates a classic photoinhibition response (Fig. S2).

Light energy absorbed in excess of what is used to fix carbon is specified by the SU rejection flux, according to:

\begin{equation} j_{eL} = (j_L - j_{CP} / y_{CL})_+ \end{equation}

This excess light energy must be quenched by alternative pathways in order to prevent photooxidative damage [@Powles:1984tm]. *Symbiodinium* utilize a variety of pathways for non-photochemical quenching [NPQ; @Roth:2014wf], which we collect in a total NPQ capacity specified as a parameter of the symbiont ($k_{NPQ}$). The NPQ flux $j_{NPQ}$ is then specified as a single-substrate SU formula with a maximum of $k_{NPQ}$:

\begin{equation} j_{NPQ} = \bigg({1 \over k_{NPQ}} + {1 \over j_{eL}}\bigg)^{-1} \end{equation}

If light energy further exceeds the capacity of both photochemistry and NPQ, then reactive oxygen species (ROS) are produced. We represent this as a relative quantity $c_{ROS}$, which takes a value of 1 when all light energy is quenched by photochemistry and NPQ, and increases as the amount of excess excitation energy increases, specified as:

\begin{equation} c_{ROS} = 1 + {{(j_{eL} - j_{NPQ})_+} \over k_{ROS}} \end{equation}

where $k_{ROS}$ is a parameter of the symbiont that determines the rate of ROS production (specifically, the amount of excess excitation energy that doubles ROS production relative to baseline levels). Importantly, $c_{ROS}$ is specified here not as a function of absolute external light, but rather the amount of excess light after accounting for quenching by carbon fixation and NPQ. A direct consequence of this formulation is that CO~2~-limitation of photosynthesis can lead to ROS production, an important mechanism [@Wooldridge:2009p7807] that was not captured by previous representations of photooxidative stress [@Eynaud:2011tv].

Carbon fixed by photosynthesis ($j_{CP}$; Eq. 14) is then combined with nitrogen shared by the host ($\rho_N$; Eq. 7) and nitrogen recycled from symbiont biomass turnover

\begin{equation} r_{NS}=\sigma_{NS}n_{NS}j_{ST}^0 \end{equation}

to build new symbiont biomass, following the SU equation:

\begin{equation} j_{SG} = \bigg({1 \over j_{SGm}} + {1 \over j_{CP}} + {1 \over {(\rho_N{H \over S} + r_{NS}) / n_{NS}}} - {1 \over {j_{CP} + (\rho_N{H \over S} + r_{NS}) / n_{NS}}}\bigg)^{-1} \end{equation}

The rejection flux of carbon from this SU represents the amount of fixed carbon produced by photosynthesis in excess of what can be used to produce symbiont biomass; this surplus, $\rho_C$, is translocated to the coral host:

\begin{equation} \rho_C = (j_{CP} - j_{SG})_+ \end{equation}

The rejection flux of nitrogen from the symbiont biomass SU is lost to the environment.

Symbiont biomass turnover includes a component of constant turnover specified by the parameter $j_{ST}^0$, representing fixed maintenance costs, plus a component that scales with the magnitude of ROS production.

\begin{equation} j_{ST} = j_{ST}^0(1 + b(c_{ROS}-1)) \end{equation}

This second component of symbiont biomass loss represents both photodamage and/or symbiont expulsion (i.e., bleaching), both of which occur in response to high levels of ROS production. The parameter $b$ is included to scale biomass loss due to bleaching in response to ROS. (Note that recycling of symbiont biomass turnover ($rNS$ and $rCS$) only occurs based on the maintenance component of turnover (i.e., $j_{ST}^0$), and not the photodamage/bleaching component, as this loss represents biomass that is damaged or expelled from the host.)

## Numerical analysis

The model dynamics are specified by the differential equations (1-2) that impose biomass balance for host and symbiont and by a set of coupled non-linear algebraic equations (3-21) that define fluxes. Several of these fluxes are defined *implicitly*; for example, the rejection fluxes of carbon and nitrogen from the symbiont and host biomass SUs, respectively, act as reciprocal input fluxes to the other SU. Similarly, the photosynthesis SU receives CO~2~ at a rate proportional to the carbon rejection flux from the host biomass SU, and the rejection flux of excitation energy from the photosynthesis SU acts to reduce its own production through photoinhibition. Without further assumptions, however, the dynamical system is not always unambiguously defined because for some combinations of parameters and environmental forcing functions the system of algebraic equations has more than one solution with all fluxes non-negative (see results below). In such circumstances, the right hand side of the differential equations (1) and (2) is not uniquely defined even when $S$ and $H$ are specified. We resolved this problem by *defining the dynamical system* as the limit as a time step $\Delta t \rightarrow 0$ of a discretized system corresponding to Euler integration of the differential equations, with those fluxes that represent flows of elemental matter implemented by assuming that transfer of material between components of the system takes one time step. Thus, for example, CO~2~ rejected from the host SU at time $t$ arrives at the photosynthesis SU at time $t + \Delta t$.

Simulations using the discretized scheme were performed using R code that is available in the data repository accompanying this article: github.com/jrcunning/Rcoral. By experimentation, we found that a time step of 0.1 days gave adequate precision for most simulations (including used to generate Figs. XXX in this paper). For steady state estimations, simulations were run until the change in specific growth rate of host and symbiont and the $S:H$ biomass ratio was less than 1e-5 per time step. In regions of state space where very slow transient dynamics could be expected (i.e. near bifurcation points), sample steady state calculations were verified using MATHEMATICA code for numerical root finding (function FindRoot) with the code written independently by a coauthor without reference to the R code.

To aid in visualizing model results, we calculated values to indicate the degree to which product formation at an SU was limited by availability of either of its two substrates using the formula

\begin{equation} \log \bigg({{\min(j_{S1}, j_{Pm})} \over {\min(j_{S2}, j_{Pm})}}\bigg) \end{equation}

where $j_{S1}$ and $j_{S2}$ are the specific input fluxes of the two substrates and $j_{Pm}$ is the maximum specific product formation rate, in units of Cmol Cmol^-1^ d^-1^.

# Steady state behavior
In a constant environment, the system ultimately reaches a steady state of exponential growth or decline. However, under some conditions, either of these outcomes may occur depending on initial values of symbiont and host biomass, indicating the presence of alternate stable states, corresponding to either positive or negative growth (Fig. 2). The mechanism that produces these alternate stable states is the positive feedback between carbon-limitation of the host and CO~2~-limitation of photosynthesis: if symbiont biomass is initially very low (i.e., a "bleached" coral), the system cannot escape this positive feedback and cannot grow. However, if symbiont biomass is initially high (i.e., a "healthy" coral), then the system remains in a nitrogen-limited state with positive growth. For practical purposes, this section of the manuscript considers only positive growth steady states under constant environments; the subsequent "dynamic behavior" section explores how changing environments may cause the system to switch between alternate stable states, which we interpret in the context of coral bleaching.

![Alternate stable states](../img/Fig2.png)

To analyze positive-growth steady state behavior, we ran the model to steady state across gradients of external irradiance and nutrients (Fig. 3), which revealed patterns consistent with observed phenomena in corals. Predicted growth rates are low at low light and DIN (~0.01 d^-1^), and begin increasing as both of these factors increase (Fig. 3A). Low light limits photosynthetic rates, resulting in less fixed carbon shared with the host and an associated increase in the symbiont to host biomass ratio (Fig. 3B). In agreement with this trend are many observations of negative correlation between irradiance and symbiont density [@Stimson:1997p3837; @Brown:1999p3534; @Fitt:2000p3792; @Titlyanov:2001p3631]. As higher light alleviates light-limitation of photosynthesis, host growth becomes less carbon-limited.

![Steady state values of **A)** specific growth (Cmol Cmol^-1^ d^-1^), **B)** the symbiont to host biomass ratio (CmolS CmolH^-1^), **C)** relative C- or N-limitation of host biomass formation, and **D)** relative CO~2~- or light-limitation of symbiont photosynthesis, across gradients of external irradiance and dissolved inorganic nitrogen. Note that typical conditions for reefs are ~1e-7 M DIN and 10-20 mol photons m^-2^ d^-1^. Simulations for each combination of light and nutrients (41 points along each axis) were run for 100 days with a time step of 1 day. Negative steady state growth rates, and corresponding S:H ratios, were set to zero.](../img/Fig3.png)

Similarly, increasing DIN alleviates nitrogen-limitation (Fig. 3A). Increased growth at higher DIN is predicted by the DEB model of @Muller:2009io, and has also been observed experimentally [@MullerParker:1994uo; @Tanaka:2007uv; @Tanaka:2013dj]. However, DIN elevation beyond a certain point (e.g., ~3-4 µM in these simulations) has little effect on growth as carbon becomes limiting. Very high nutrient levels may even reduce growth [@Shantz:2015dz], although these impacts are not likely to occur within the range of concentrations considered here (<4 µM) [@FerrierPages:2000p1950]. In addition to increasing growth, DIN also increases the symbiont to host biomass ratio (Fig. 3B), a phenomenon also observed in reef corals [@Marubini:1996p8460]. At low DIN and intermediate light, more typical of coral reef environments, symbiont to host biomass ratios are around ~0.06-0.21, which is consistent with values reported in the literature [@Muscatine:1981jy; @Edmunds:2011bv; @Hawkins:2016jj].

The maximum predicted growth rates of ~0.1 d^-1^, occurring between ~10-25 mol photons m^-2^ s^-1^ light and ~4 µM DIN (Fig. 3A), are comparable to the rate of `r Tanaka_gr` d^-1^ measured by @Tanaka:2007uv under similar N-enriched conditions. Predicted growth rates under conditions more typical of reef environments (<0.5 µM DIN) are ~0.01-0.03 d^-1^. Observed specific growth rates, which are typically based on skeletal mass, fall near or below 0.01 d^-1^ [@Osinga:2011kq; @Osinga:2012is], though values as high as 0.025 d^-1^ have been reported [@Schutter:2010p7758]. It is not surprising that predicted growth rates are slightly higher than most observations, as the model does not account for external ecological factors that may reduce growth, such as competition, predation, and bioerosion. Furthermore, the model does not explicitly account for skeletal formation, which may not always correlate with biomass growth, potentially due to differential substrate-limitation [@Anthony:2002tc]. Nevertheless, when growth of coral tissue biomass has been directly measured over short periods, values are consistent with model predictions (e.g., `r Tanaka_gr` d^-1^ in @Tanaka:2007uv), and similar rates (`r Armoza_gr` d^-1^) have been measured in *Aiptasia diaphana*, a non-calcifying symbiotic anemone [@ArmozaZvuloni:2014ju].

As irradiance continues to increase above ~25 mol photons m^-2^ d^-1^, growth rates decline until positive growth ceases above ~40 µmol photons m^-2^ d^-1^ (Fig. 3A). The mechanism underlying this decline is the increase in light energy beyond the capacities of photosynthesis and non-photochemical quenching: excess excitation energy generates reactive oxygen species (ROS) [@Weis:2008p944; @Roth:2014wf], which, in this model, have the phenomenological consequences of reducing the photosynthetic rate (representing photoinhibition) and increasing symbiont biomass loss (representing photodamage and/or symbiont expulsion) [see @Eynaud:2011tv]. Together, these impacts reduce the symbiont to host biomass ratio (Fig. 3B), as occurs during coral bleaching. This reduction in symbionts consequently reduces the flux of fixed carbon to the host, resulting in increasing carbon-limitation (Fig. 3B) and eventual cessation of growth (Fig. 3A).

The incorporation of photooxidative stress in the model sets an upper limit to the amount of light at which a stable symbiotic interaction can be maintained, but even below this threshold of breakdown, negative effects of high light reduce steady state growth and symbiont:host biomass (Fig. 3). This gradual decline is consistent with experimental results showing that high light levels decrease growth [@Schutter:2011ie], and field studies documenting optimum growth rates at intermediate depths [@Baker:1975ip; @Huston:1985ec]. By incorporating these impacts of light stress, the model predicts greater, and more realistic, variation in state variables across light gradients than was predicted by the model of @Muller:2009io, which did not include photoinhibition and photodamage. It is important to recognize that the boundary set by photooxidative stress on a stable symbiosis under steady state conditions (Fig. 3) may be temporarily crossed by a dynamic system, which may experience a period of symbiont loss (bleaching) and reduced growth, after which a return to benign conditions may restore symbiont biomass and positive growth. To explore this further and illustrate the behavior of the model in more detail, we evaluate a number of dynamic simulations below (see "Dynamic behavior").

# Sensitivity analysis

The values used for each parameter in the model (Table 2) are derived from relevant literature; these derivations are described in the Supplementary Information. Here we evaluate the sensitivity of the model to changes in these parameter values, which also serves to demonstrate the behavior of the dynamical system. We measured fractional change in steady state values in response to fractional changes in parameter values, relative to their default values, under environmental conditions typical of coral reefs.

![Sensitivity analysis. Plots show the fractional change in steady state values of growth (solid lines) and S:H biomass (dashed lines) in response to fractional changes in default parameter values (see Table 2 for default values). Parameters are grouped by which processes they are involved in. This sensitivity analysis was conducted at conditions typical for coral reef environments: low DIN (1e-7 M) and intermediate light (15 mol m^-2^ d^-2^). Sensitivity analyses conducted other environmental conditions are presented in Figs S3-S7.](../img/Fig4.png)

Overall, relative changes in the steady state of the system are less than the equivalent relative change in parameter value, indicating that the system maintains internal stability. However, changes in certain parameter values have more significant impacts: increasing $j_{Nm}$ or decreasing $K_N$ both dramatically increase host growth (Fig. 4), demonstrating the strong nitrogen-limitation that characterizes these symbioses. The parameter $\bar{a}^*$ has a strong impact on S:H biomass ratios (Fig. 4) since this parameter determines the amount of light absorbed by symbionts, with lower values increasing light-limitation. Increasing the maximum growth and turnover rates have the expected effects of increasing and decreasing growth, respectively. Parameters relating to photooxidative stress and bleaching have little impact under these environmental conditions (Fig. 4), but have larger impacts under higher light (e.g., Fig. S6). Sensitivity analyses conducted under different combinations of external light and nutrients are presented in Figs. S3-S7.

# Dynamic behavior
The dynamic behavior of the model demonstrates its power to integrate multiple environmental forcings simultaneously. Here we present several scenarios that demonstrate the model's ability to reproduce complex phenomena that have been observed in corals.

## Seasonal variability
Symbiont densities and coral growth rates are known to vary seasonally, representing an integrated response to changes in a suite of environmental factors. Light in particular is a strong driver of these trends [@Stimson:1997p3837; @Brown:1999p3534; @Fagoonee:1999p4136; @Fitt:2000p3792], with high light associated with lower symbiont abundance and reduced tissue biomass. The role of light in driving seasonal changes in symbiont density was demonstrated nicely by @Stimson:1997p3837, who also found that experimental nutrient-enrichment amplified the light-driven seasonal oscillation. Using the levels of light and nutrients from this study as inputs, the model reproduces this observed interaction among environmental factors (Fig. 5), and also provides the mechanism: under high light in summer, photooxidative stress decreases symbiont biomass and increases carbon-limitation of host growth regardless of nutrient status; when light is reduced in winter, growth and symbiont biomass become constrained instead by nitrogen, and therefore increase more when DIN is enriched (Fig. S8).

![Light-driven seasonal dynamics of symbiont abundance and coral growth. Light input (upper panel) was designed as a sinusoidal curve with a period of one year, with maximum and minimum values of 44 and 22 mol photons m^-2^ d^-1^, based on @Stimson:1997p3837. The dynamic behavior of symiont to host biomass ratio (middle panel) and the specific growth rate of host biomass (lower panel) show seasonal oscillations that are greater in magnitude under higher nutrients (1e-5 molN L^-1^; dashed lines) relative to lower nutrients (1e-7 molN L^-1^; solid lines), consistent with the findings of @Stimson:1997p3837. Prey density was set at 1e-5 CmolX L^-1^.](../img/Fig5.png)

The prediction of higher growth when light is reduced indicates that growth is not limited by low light in winter, but is actually reduced by excess light in summer, consistent with the experimental findings of @Schutter:2011ie. Seasonal summertime reductions in tissue biomass have also been well-documented in the field [@Fitt:2000p3792], along with reductions in net photosynthetic capacity [@MullerParker:1987cr]. Importantly, while light alone (at these levels) may drive seasonal dynamics in the ways discussed, temperature fluctuations may attenuate or even reverse the effect of light as cooler winters depress metabolism; thus, the relative magnitude of fluctuation in temperature, light, and other factors may produce wide variability in the direction and magnitude of seasonal changes in growth and symbiont abundance, depending on location and microhabitat. Nevertheless, the seasonal variability predicted here (Fig. 5) is consistent with experimental and field observations for corals, and demonstrates the power of this model to predict dynamic behavior that mechanistically integrates multiple environmental drivers.

## Coral bleaching and recovery
Coral bleaching is the stress-induced loss of symbiotic algae from coral tissues, which can occur in response to a variety of environmental stressors. In most cases, coral bleaching is thought to begin with photooxidative stress in symbiont photosynthesis [@Lesser:1997p3587], which triggers a cascade of events leading to symbiont expulsion [@Weis:2008p944]. As symbionts are expelled, the host receives less fixed carbon, which may then compromise its ability to activate CCMs that deliver CO~2~ to photosynthesis [@Wooldridge:2013tj]. Increasing CO~2~-limitation for remaining symbionts, along with an amplified internal light environment due to reduced self-shading [@Enriquez:2005p142], may further exacerbate photooxidative stress and accelerate symbiont expulsion, driving the coral into a bleached state.

While these positive feedbacks have been discussed previously in the literature, this is the first attempt to implement and explore their properties within a dynamical model. Interestingly, these feedbacks lead to alternate stable states in the symbiotic system. The 'healthy' state is characterized by nitrogen-limitation of both symbiont and host: under these conditions, the symbiont translocates sufficient carbon to support host growth and CCMs, which ensures that photosynthesis does not become CO~2~-limited. However, if carbon translocation is disrupted (and light is sufficiently high), then the system is driven into the 'bleached' state by photooxidative stress and positively reinforcing carbon- and CO~2~-limitation. In this context, coral bleaching can be understood as a transition from one stable state to another, and bleaching thresholds are sets of environmental conditions that push a healthy-state coral onto a trajectory leading to a bleached state.

We are highly interested in the conditions under which the system switches from a healthy to a bleached state, and can use this model as a tool to explore this dynamic behavior. Most straightforwardly, this switch occurs when increasing external irradiance (Fig. 6A) causes sufficient ROS production (Fig. 6B) and photoinhibition (Fig. 6C) such that the positive feedbacks between host carbon-limitation (Fig. 6D) and CO~2~-limitation of photosynthesis (Fig. 6E) are rapidly engaged, leading to even greater photooxidative stress and a rapid decline in S:H biomass (Fig. 6F), characteristic of coral bleaching. However, the positive feedbacks involved in bleaching are not engaged only in response to high external irradiance alone; in fact, they depend on the relative balance of light energy absorption and quenching, which in turn depends on the availability of CO~2~ for photosynthesis. While previous models framed photooxidative stress as a fixed response to absolute external irradiance [@Eynaud:2011tv], our implementation considers the dynamic balance of multiple energy sinks in the causation of stress, which is more consistent with current understanding of symbiosis dysfunction [@Wooldridge:2013tj], and establishes a critical role of host CCMs in providing CO~2~ for photosynthesis [@Tansik:2015cm; @Hopkinson:2015ev].

![Coral bleaching as a switch from a nitrogen- to carbon-limited alternative stable state. This transition is demonstrated in response to gradually increasing external light (A), which causes an increase in production of ROS (B) that reduces the photosynthetic rate through photoinhibition (C). Decreasing photosynthesis moves the system from overall nitrogen-limitation toward carbon-limitation (D); when this threshold is crossed, the system rapidly becomes highly carbon-limited as photosynthesis becomes CO~2~-limited (E) and symbiont densities rapidly decline (F) into a bleached state. (All parameters at default values; external DIN=1e-7 molN/L, prey density=0).](../img/Fig6.png)

The importance of host CCM activity establishes significant interactive roles for other factors in influencing coral bleaching responses. For example, simulations of high light stress (Fig. 7) demonstrate that bleaching can be attenuated by heterotrophic feeding, a phenomenon which has been observed experimentally [@Borell:2008p108]. The mechanism underlying this prediction is that feeding by the host increases host CCM activity, which delays the onset of CO~2~-limitation of photosynthesis and reduces bleaching severity. On the other hand, elevated nutrients exacerbate bleaching (Fig. 7), since higher symbiont densities are more susceptible to CO~2~-limitation [@Wooldridge:2009p7807]. Several experimental [@Cunning:2013gp; @VegaThurber:2014hk] and correlational studies [@Wooldridge:2009kb] are consistent with this mechanistic link between high nutrients and bleaching.

![Bleaching with interactive factors. Simulations of high light stress (sinusoid with maximum=48 mol ph m^-2^ d^-1^) under default environmental conditions (solid line; DIN=1e-7 molN/L; prey=3e-6 CmolX/L), or with elevated feeding (dashed line; prey=6e-6 CmolX/L), or elevated nutrients (dotted line; DIN=1e-6 molN/L). Initial symbiont biomass was set to the steady state for each set of starting conditions, respectively. (All parameters set to default values).](../img/Fig7.png)

Since bleaching is induced by an external stressor, recovery from bleaching cannot occur unless the stressor is alleviated. In natural settings, this typically occurs by seasonal declines in temperature and light. However, alternate stable states lead to significant hysteresis, indicating that a coral cannot recover along the same trajectory it followed during bleaching; indeed, the stressor must be alleviated well *below* the threshold that initially caused bleaching in order for the system to recover (Fig. 2; Fig. X). This is because under the same external conditions, a bleached coral (relative to a healthy one) is characterized by greater light amplification and weaker CCM activity, which serve to maintain the bleached state. The magnitude of hysteresis will depend on external nutrients and prey availability, as well as model parameters, yet the core behavior of hysteresis illustrates the alternate stable states that exist in the symbiotic system.

![Recovery from bleaching under varying N:C availability. Higher N:C ratios in a heterotrophic food source (effectively representing higher external DIN and/or lower feeding rates) cause a larger overshoot in the S:H biomass ratio, and prolong the duration of time until the system 'recovers' by re-establishing nitrogen-limitation (represented by black dots on each trajectory). All simulations run with default parameters (except varying nNX), and DIN=1e-7 molN/L, prey=1e-6 CmolX/L, and initial symbiont biomass=0.0001.](../img/Fig8.png)

The dynamics of recovery from bleaching reveal another interesting behavior of the system: under some conditions, the S:H biomass ratio may increase to a higher value that 'overshoots' the ratio maintained in the 'healthy' state under the same conditions (Fig. 8; Fig. S9). In fact, unusually high symbiont densities have been observed after recovery from bleaching in both experimental [@Cunning:2015ja] and field studies [@Kemp:2014ug], and have been interpreted as a potential 'disequilibrium in host-symbiont regulation' [@Kemp:2014ug]. The model explains this 'overshoot' as an approach to a bifurcation point in the dynamical system <<NOT RIGHT>>. As the symbiont begins repopulating the host, CO~2~ is limiting and nitrogen is replete, allowing nearly all fixed carbon to be used for symbiont growth. As a result, host growth remains strongly carbon-limited, CCM activity remains weak, and photosynthesis remains CO~2~-limited. As long as this carbon-limitation persists, S:H biomass will continue to increase beyond the 'healthy' steady state level, until the symbiont population eventually reaches a high enough density such that nitrogen limits its growth more strongly than carbon. At this point, which represents the peak of the 'overshoot', the symbiont begins to share carbon with the host, which then supplies more CO~2~ for photosynthesis, etc., causing the system to rapidly escape the positive feedbacks that were maintaining carbon-limitation (see Fig. S9 for illustration of all fluxes during overshoot). The re-establishment of nitrogen-limitation resumes high rates of carbon fixation and coral growth, bringing the S:H biomass ratio back down to a stable equilibrium.

The occurrence and the magnitude of this overshoot depend on the relative inputs of carbon and nitrogen to the system -- any factor that enhances carbon-limitation (e.g. high DIN and/or low feeding) therefore magnifies the overshoot and prolongs the dysfunctional, carbon-limited state of the symbiosis (Fig. 8). On the other hand, factors that favor nitrogen-limitation, such as low external DIN, high feeding rates, will prevent an overshoot from occurring at all, since nitrogen-limitation is re-established in the system before S:H biomass ratios reach high levels. While many scenarios are possible under different environmental conditions, we illustrate the general effects of varying N and C availability on recovery from bleaching with a series of simulations that vary the N:C ratio of host's heterotrophic food source (Fig. 8): lower N:C ratios (effectively representing lower DIN and/or higher heterotrophy) favor nitrogen-limitation and more rapid recovery, while higher N:C ratios (effectively representing higher DIN and/or lower heterotrophy) favor carbon-limitation and prolonged recovery with a larger overshoot in S:H biomass.

Importantly, the re-establishment of nitrogen-limitation that characterizes the 'healthy' symbiosis is the most important diagnostic for recovery, as this is when high carbon fixation and growth rates are resumed. A high S:H biomass ratio alone does not indicate a 'healthy' state has been reached, because overall C-limitation could still persist and thus the host isn't really growing. This could explain why corals that have apparently recovered their symbiont populations after bleaching still do not grow or suffer reduced reproduction until much later (xxx). Together, these dynamics indicate that acquisition of carbon from a source other than the symbiont is extremely important for the host in order to recover from bleaching. The system can 'recover' (in terms of resuming positive growth and N-limitation) more rapidly when the coral obtains more carbon. Indeed, feeding has been shown to promote more rapid recovery from bleaching (xxx). Moreover, if the host could obtain an independent source of carbon, such as direct uptake of DOC (xxx), this could be even more important in ensuring rapid recovery from a bleached state.

# Conclusions
This simplified dynamic bioenergetic model of coral-*Symbiodinium* symbioses mechanistically reproduces patterns in steady-state coral growth and symbiont abundance commonly observed in corals, including higher symbiont abundance with higher nutrients and feeding, lower symbiont abundance with increasing light, and optimal growth at intermediate light levels. Moreover, the model reproduces complex dynamic behaviors including seasonal changes in symbiont density at different nutrient levels, rapid bleaching above a threshold of high light, mitigation of bleaching by heterotrophic feeding, exacerbation of bleaching by elevated nutrients, and an overshoot of symbiont density during recovery from bleaching. These examples demonstrate the model's ability to integrate multiple environmental forcing functions to reproduce complex responses; meanwhile, the diversity of these phenomena suggest the model has captured many of the important features of the system in a unifying mechanistic framework. This model also provides a new conceptual framework for considering coral bleaching as a transition to an alternate stable state, which has important implications for understanding the performance and maintenance of symbiotic interactions. In this context, the 'healthy' stable state represents a scenario in which the host is 'in control' of the system, with nitrogen-limitation stabilizing the symbiont to host biomass ratio and maintaining positive growth. Conversely, carbon-limitation represents a loss of host 'control', wherein positive feedbacks result in a loss of capacity for photosynthetic carbon fixation and positive growth. Importantly, these 'control' mechanisms arise entirely passively through feedbacks present in the system.

In addition to the suite of specific examples presented here, the model can be used to explore many different dynamic environmental scenarios, and represents a tool that biologists and ecologists can use to generate hypotheses and make predictions in both experimental and natural settings. Moreover, parameter values can be modified to correspond to different genetic or functional types of coral hosts and *Symbiodinium* partners in order to evaluate variability in system responses. Thus, the diversity of potential applications for this model is high, and we envision this work as a foundation for continued development, which may include more detailed treatments of specific modules (e.g., DIC processing), and the incorporation of more external forcing capabilities (e.g., external DIC). Importantly, this effort is best undertaken by the scientific community at large, including those with theoretical and empirical backgrounds, which is why the code for this model has been developed openly using the R language. Ultimately, the continued refinement of these tools is fundamental in elucidating the mechanisms of symbiosis function and dysfunction, and in predicting coral responses to environmental change.

References {#references .unnumbered}
==========